Rusty Linux: Advances in Rust for Linux Kernel Development

### 2024-10-24 slides.tex
Final updates to slides with ai gen pics

\subsection{Research Questions}
\begin{frame}
  \frametitle{Research Questions}
  \begin{itemize}
  \item<1-> \textbf{RQ1:} What are the existing approaches for implementing operating system kernels in Rust?
  \item<2-> \textbf{RQ2:} What are the performance implications of using Rust for operating system kernel development?
  \item<3-> \textbf{RQ3:} What are the major challenges and limitations when developing operating system kernels in Rust?
  \item<4-> \textbf{RQ4:} What are the lessons learned when developing operating systems kernels in Rust?

  \end{itemize}
\end{frame}

@@

\subsection{RQ3: Challenges and Limitations}
\begin{frame}
  \frametitle{RQ3: Challenges and Limitations}
  \begin{itemize}
  \item<1-> Binary Size
    \begin{itemize}
    \item Rust can produce larger binaries
    \item The same issue that C++ templates have!
    \end{itemize}
  \item<2-> Missing Features
    \begin{itemize}
    \item Rust still evolving and adding features
    \item Makes it difficult to integrate into the classroom due to the rapid evolution
    \end{itemize}
  \item<3-> Soundness - How to deal with raw memory?
  \item<4-> Panics - What happens when things go wrong?
  \item<5-> C Interop - Specific to mixed language kernels
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{RQ4: Lessons Learned}
  \begin{itemize}
  \item<1-> Impossible to use 100$\%$ rust
  \item<2-> Rust is not as expressive as other formal verification techniques
  \item<3-> Ownership root - An OS provides memory to rust so if the OS is itself written in rust
    who is the root owner?
    \begin{itemize}
    \item Open research question if this can even be done in software
    \item Researchers looking at hardware support (CHERI)
    \end{itemize}
  \end{itemize}

>>>>

\subsection{Research Questions}

\subsubsection{RQ1}
\begin{frame}
  \frametitle{RQ1}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \textbf{RQ1:} What are the existing approaches for implementing operating system kernels in Rust?\footnotemark[2]
    \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[width=.8\textwidth]{images/rq1.jpeg}
    \end{column}
  \end{columns}
  \footnotetext[2]{Fun side note: What happens if we put our research questions into an AI image generator?}
\end{frame}

\subsubsection{RQ2}
\begin{frame}
  \frametitle{RQ2}
    \begin{columns}
      \begin{column}{0.5\textwidth}
        \textbf{RQ2:} What are the performance implications of using Rust for operating system kernel development?
    \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[width=.8\textwidth]{images/rq2.jpeg}
    \end{column}
  \end{columns}
\end{frame}

\subsubsection{RQ3}
\begin{frame}
  \frametitle{RQ3}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \textbf{RQ3:} What are the major challenges and limitations when developing operating system kernels in Rust?
    \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[width=.8\textwidth]{images/rq3.jpeg}
    \end{column}
  \end{columns}
\end{frame}

\subsubsection{RQ4}
\begin{frame}
  \frametitle{RQ4}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \textbf{RQ4:} What are the lessons learned when developing operating systems kernels in Rust?
      \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[width=.8\textwidth]{images/rq4.jpeg}
    \end{column}
  \end{columns}
\end{frame}

@@

\subsection{RQ3: Challenges and Limitations}
\begin{frame}
  \frametitle{RQ3: Challenges and Limitations}
  \begin{itemize}
  \item<1-> Binary Size
    \begin{itemize}
    \item Rust can produce larger binaries
    \item The same issue that C++ templates have!
    \end{itemize}
  \item<2-> Missing Features
    \begin{itemize}
    \item Rust still evolving and adding features
    \item Makes it difficult to integrate into the classroom due to the rapid evolution
    \end{itemize}
  \item<3-> Soundness - How to deal with raw memory without sacrificing safety?
  \item<4-> Panics - What happens when things go wrong?
  \item<5-> C Interop - Specific to mixed language kernels
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{RQ4: Lessons Learned}
  \begin{itemize}
  \item<1-> Impossible to use 100$\%$ rust - Same with C, some low level asm is needed to setup initial stack
    pointer, etc.
  \item<2-> Rust is not as expressive as other formal verification techniques
  \item<3-> Ownership root - An OS provides memory to rust so if the OS is itself written in rust
    who is the root owner?
    \begin{itemize}
    \item Open research question if this can even be done in software
    \item Researchers looking at hardware support (CHERI)
    \end{itemize}
  \end{itemize}
